Robust calibration of numerical models based on relative regret

### 2019-11-19 article-modif-A&E.tex
fix space in front of cite, booktabs

%% For including figures, graphicx.sty has been loxaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

@@

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
\usepackage{easy-todo}
\usepackage{graphicx}

% À changer
\newcommand{\yobs}{\mathbf{y}^o}
\DeclareMathOperator*{\argmin}{arg\,min \,}
\DeclareMathOperator*{\argmax}{arg\,max \,}
\newcommand{\Var}{\mathbb{V}\textrm{ar}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Cov}{\textsf{Cov}}
\newcommand{\tra}{\mathrm{tr}}
\newcommand{\kmean}{{\mathbf{k}}_{\mathrm{mean}}}
\newcommand{\hatkmean}{\hat{\mathbf{k}}_{\mathrm{mean}}}
\newcommand{\kvar}{{\mathbf{k}}_{\mathrm{var}}}
\newcommand{\hatkvar}{\hat{\mathbf{k}}_{\mathrm{var}}}
\newcommand{\kmpe}{{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\hatkmpe}{\hat{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\kest}{\hat{\mathbf{k}}}

\newcommand{\checkap}{{\alpha}_p}
\newcommand{\checka}{{\alpha}}
\newcommand{\checkk}{\mathbf{k}}
\newcommand{\checkkp}{{\mathbf{k}}_p}
\newcommand{\Kspace}{\mathbb{K}}
\newcommand{\Uspace}{\mathbb{U}}

\newcommand{\arthur}[1]{{\itshape \color{cyan} ({#1})}}
\newcommand{\elise}[1]{{\itshape \color{red} ({#1})}}
\newcommand{\victor}[1]{{\itshape \color{green} ({#1})}}

@@

Numerical models are widely used to study or forecast natural phenomenon and improve industrial processes. However, by essence models only partially represents reality and sources of uncertainties are ubiquitous (discretisation errors, missing physical processes, poorly known boundary conditions).
%Many physical phenomena are modelled numerically in order to better understand and/or to predict their behaviour. Because of the complexity and or ignorance of the underlying physics, numerical models are imperfect representations of the reality. 
%
%Various sources of uncertainties can corrupt the model quality, such as the discretisation of the domain, or presence of  small scale phenomena that can not be fully represented. 
Moreover, such uncertainties may be of different nature. 
 \cite{walker_defining_2003} propose two to consider two categories of uncertainties. On the one hand aleatoric uncertainties,  coming from the inherent variability of a phenomenon, e.g. intrinsic randomness of some environmental variables. On the other hand, epistemic uncertainties coming from a lack of knowledge about the properties and conditions of the phenomena underlying the behaviour of the system under study.
The latter can be accounted for through the introduction of ad-hoc correcting terms in the numerical model, that need to be properly estimated. Thus, reducing the epistemic uncertainty can be done through parameters estimation approaches. 
%when one is having observations of the reality. 
This is usually done using optimal control techniques, leading to an optimisation of a well chosen cost function which is typically built as a comparison with reference observations.
%
An application of such an approach, in the context of ocean circulation modeling, is the estimation of ocean bottom friction parameters in \cite{das_estimation_1991} and \cite{boutet_estimation_2015}.

@@

Accounting for the probabilistic nature of  $\mathbf{U}$ leads to study the distribution of the random variable $J(\mathbf{k},\mathbf{U})$, or the distribution of its minimisers ${\mathbf{k}}_{\mathbf{U}}$. The latter is referred as 
the distribution of the conditional minimisers, notion that appeared notably in~\cite{villemonteix_informational_2006} and in~\cite{hennig_entropy_2011} for a global optimisation purpose. Both approaches and related robust estimates are described in section \ref{robust formulations}. Section \ref{sec:relax_constraint} introduces a new class of estimators, by relaxing the constraint of optimality and defining regions of acceptability, similarly as~\cite{buhmann_robust_2013} in discrete combinatorial problems.  The intention is that a robust estimate provides values of the cost function close enough to the attainable minimum for each configuration induced by $\mathbf{u}\in\Uspace$. Illustration of the various discribed methods are given on a numerical exemple in section \ref{sec:SWE_application}.

@@

As mentioned before, robustness can be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$. Under this definition, one may design classical robust estimators, either by optimising the moments of the cost function ; or based on the distribution of its minimisers.

@@

As the computation of this Pareto front is usually hard and expensive, alternative strategies based on the minimisation of a scalarized version of the vector of objectives are often considered. Some are based on a weighted sum of the objectives, as presented in~\cite{grodzevich_normalization_2006} and in \cite{marler_weighted_2010}, while some others are based on the minimisation of one of the objectives under constraints on the others, as performed in~\cite{lehman_designing_2004}. Both of these methods are based on an \textit{delicate} choice of weights or of constraints before any computation. This choice relies heavily on a knowledge of the properties of the cost function.

@@

Once again, a set of samples  $\{\mathbf{u}^i\}_{i=1\dots N}$ can be used to compute the set $\{\mathbf{k}_{\mathbf{u}^i}\}_{i=1\dots N}$, from which one can 
approximate  $p_{\mathbf{K}^*}$. The resulting approximation and therefore its mode, is sensitive to the density estimation method. Main methods are KDE  (Kernel Density Estimation) \cite{elise}, and  EM (Expectation-Maximisation)~\citep{dempster_maximum_1977}.

@@

\begin{figure*}[!ht]
  \includegraphics[width=12cm]{Figures/branin_side_moments_noscale.pdf}
  \caption{Left: Contour of the $J_{\mathrm{BH}}$ function, conditional mean and variance. The estimates $\hatkmean$ and $\hatkvar$ are plotted with the dashed line. Right: $J_{\mathrm{BHswap}}$ function. The scale of the bottom plots have not been represented.} %TODO revoir
\label{fig:branin_moments}
\end{figure*}
%\clearpage % Pour lisibilité dans le preview

@@

Illustrations related to $\hatkmpe$ are depicted on Figure \ref{fig:contours}. As before, the set of $\mathbf{u}^i$ is taken from a regular grid. Top row shows the contour plot of both functions as well as the set of conditional minimisers $\{\mathbf{k}_{\mathbf{u}^i}\}_{1\leq i \leq N}$ in red, as defined in Eq.~\eqref{eq:conditional_minimiser}.
The left, respectively right, column stands for $J_{\mathrm{BH}}$, respectively $J_{\mathrm{BHswap}}$ function. The bottom rows presents three approximations of the density of minimisers: the histogram in grey (bin size selected using Freedman-Diaconis from \cite{freedman_histogram_1981}), the result of a kernel density estimation (KDE) with Gaussian kernels in red (using Scott's rule from \cite{scott_optimal_1979} for bandwidth selection), and the estimation by a Gaussian mixture. The latter has been calculated with the EM algorithm. The number of Gaussians has been fixed to 3, a guess based on the general shape of the histogram. Respective estimations of $\hatkmpe$ are also depicted using dashed lines.

@@

\begin{figure*}
  \includegraphics[width=12cm]{Figures/branin_side_66.pdf}
  \caption{Top Left:  $J_{\mathrm{BH}}$ function along with conditional minimisers in red. Bottom left: Estimated density using KDE and EM algorithm and conditional mean. The dashed lines indicate the MPE found using those methods. Right: Same quantities with $J_{\mathrm{BHswap}}$. The estimation of the MPE is consistent between the methods for $J_{\mathrm{BHswap}}$, but not for  $J_{\mathrm{BH}}$}
\label{fig:contours}
\end{figure*}

@@

\begin{figure*}[!t]
\includegraphics[width=12cm]{Figures/relaxation_tuto.pdf}
\caption{Principle of the relaxation of the constraint on $J_{\mathrm{BHswap}}$, and illustration of $R_\alpha(\mathbf{k})$}
\label{fig:relax_tuto}
\end{figure*}

%In order to construct this region, we introduce a factor $\alpha \geq 1$, that will be multiplied to $J^*$:
%for a given $\mathbf{u}$, $\mathbf{k}$ is deemed acceptable if $J(\mathbf{k},\mathbf{u}) \leq \alpha J^*( \mathbf{u})$.
Introducing the random nature of $\mathbf{U}$, $\Gamma_\alpha(\mathbf{k})$ is defined as the probability that $\mathbf{k}$ is acceptable given $\alpha$:
\begin{equation}
  \label{eq:def_Gamma}
  \Gamma_\alpha(\mathbf{k}) = \Prob_{\mathbf{U}}\left[\mathbf{U} \in R_\alpha(\mathbf{k})\right] = \Prob_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U}) \leq 
    \alpha J^*(\mathbf{U}) \right]
\end{equation}
In other words, $\Gamma_{\alpha}(\mathbf{k})$ is then the probability that $J(\mathbf{k},\mathbf{U})$ is between 
$J^*(\mathbf{U})$ and 
$\alpha J^*(\mathbf{U})$. Without relaxation, i.e.\ when $\alpha$ is set to $1$, $\Gamma_1$ is non-zero if the set $\{\mathbf{u}\in\Uspace \mid J(\mathbf{k},\mathbf{u}) = J^*(\mathbf{u})\}$ has non-zero measure with respect to $\Prob_{\mathbf{U}}$. Still, this can be related to the definition of the MPE from Eq.~\eqref{eq:MPE}

@@

Given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$. This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see \cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. The associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$.

\begin{figure}[!ht]
\centering
\includegraphics[width = 8.3cm]{Figures/illu_alpha_p.pdf}
\caption{Illustration of the influence of different levels $p$ on $\Gamma_{\alpha_p}$ and on $\mathbf{k}_p$ \victor{Est-ce que tu pensais à ce genre de figure ?}}
\label{fig:illu_alpha_p}
\end{figure}

@@

\begin{table*}[t]
  \centering
  \caption{Robust estimators, based on a cost function $J$}
  \label{tab:RO_recap}
\begin{tabular}{lrr}
  \hline
  Definition & Related quantities & Interpretation \\
  \hline
   $\argmin_{\mathbf{k}\in\Kspace} \Ex_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kmean$ & Long run performances\\
   $\argmin_{\mathbf{k}\in\Kspace} \Var_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kvar$  & Steady performances\\
   $\argmax_{\mathbf{k}\in\Kspace} p_{\mathbf{K}^*}(\mathbf{k})$ & $\kmpe$ & Most probable minimiser\\
    $\inf\left\{ \alpha \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\}$ & $(p, \checkkp,\checkap)$ & Acceptable values with fixed probability $p$ \\
    $p_{\mathrm{maxratio}} = \argmax p/\checkap$ & $(p_{\mathrm{maxratio}}, \checkk_{\mathrm{maxratio}}, \checka_{\mathrm{maxratio}})$ & Maximal ratio of $p$ and $\checkap$ \\
  \hline
\end{tabular}
\end{table*}

@@

% ----------------------------------------------------------------------------
\subsection{Estimation of $\checkkp$ and $\checkap$ for the $J_{\mathrm{BH}}$ and $J_{\mathrm{BHswap}}$ functions}
%Numerical application}
As stated before, we chose to model the uncertainties as a random variable uniformly distributed on $\Uspace$. The bounded nature of $\Uspace$ allow us to consider the estimate introduced before up to a level of confidence $p=1$.
From now on, $\hatkmpe$ is estimated using KDE with Gaussian kernels.
The smallest relaxation $\hat{\checka}_1$ and the corresponding $\hat{\checkk}_1$ has been computed for the $J_{\mathrm{BH}}$ and $J_{\mathrm{BHswap}}$ functions, using a regular grid of $1000 \times 1000$ points on $\Kspace\times\Uspace$. The contour plots of those functions can be seen in the top plots of Figure~\ref{fig:contour_alpha}. The frontier corresponding to the couples of points $(\mathbf{k},\mathbf{u})$ verifying $\{J(\mathbf{k},\mathbf{u}) = \hat{\checka}_1 J^*(\mathbf{u})\}$ has been drawn on top of these contour plots. 
On the bottom plots, the curve $\mathbf{k}\mapsto \hat{\Gamma}_{\alpha}(\mathbf{k})$ for $\alpha=\hat{\checka}_1$ and $\alpha=1.5$ along with the histogram of the minimisers have been represented.

@@

\begin{table*}[!ht]
\centering
\caption{Estimation performed for $J_{\mathrm{BH}}$, sorted by value}
\label{tab:recap_estimates_branin}
\begin{tabular}{lr}
\hline
Estimate & Value \\ \hline
$\hatkvar$ & 1.371 \\ 
$\kest_p,~p=1$ & 1.557 \\ 
$\hatkmean$ & 1.587 \\ 
  $\hatkmpe$ & 1.628 \\
  $\kest_{\mathrm{maxratio}},~\hat{p}_{\mathrm{maxratio}}=0.654$ & 1.637 \\
$\kest_p,~p=0.95$ & 1.672 \\ \hline
\end{tabular}
\end{table*}

Practically speaking, in order to compare the effective values taken by the objective function given an estimate $\kest$ we are going to consider the functions of the form $\mathbf{u} \mapsto J(\kest,\mathbf{u})$, that we will call "profile of $\kest$". Those profiles are well suited for the representation of the cost function for an estimate $\kest$ fixed as the uncertain variable is modelled with a 1D uniform random variable.

@@

\begin{table*}[!h]
\centering
\caption{Estimations performed for $J_{\mathrm{BHswap}}$, sorted by value}
\label{tab:recap_estimates_BHs}
\begin{tabular}{lr} \hline
Estimate & Value \\ \hline
$\kest_{\mathrm{MPE}}$ & 0.606 \\ 
$\kest_{\mathrm{maxratio}},~ \hat{p}_{\mathrm{maxratio}}=0.766$ & 1.537 \\ 
$\kest_{\mathrm{mean}}$ & 1.752 \\ 
$\kest_p,~p=0.90$ & 2.112 \\ 
$\kest_p,~p=0.95$ & 2.457 \\ 
$\kest_{\mathrm{var}}$ & 2.638 \\ 
$\kest_1,~p=1$ & 2.798   \\ \hline
\end{tabular}
\end{table*}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=.7\textwidth]{Figures/RobustSWEmis_plots.pdf}
\caption{Procedure of robust calibration for the shallow water problem. Top: contours of $J$, conditional minimisers and $\{(\mathbf{k},\mathbf{u}) | J(\mathbf{k},\mathbf{u}) = \hat{\alpha}_1 J^*(\mathbf{u})\}$. Middle: Conditional moments and histogram and KDE of the conditional minimisers. Bottom: $\hat{\Gamma}_{\alpha_p}$ for different levels $p$}
\label{fig:estimation_swe}
\end{figure}
\begin{table*}[!h]
\centering
\caption{Calibrated values of $\mathbf{k}$ according to different criteria}
\begin{tabular}{lr} \hline
Estimator & Value \\ \hline
$\hatkmpe$ & 0.249 \\
$\kest_{\mathrm{global}}$ & 0.290 \\
$\hat{\mathbf{k}}_{p},~p=1$ & 0.423 \\ 
$\hat{\mathbf{k}}_p,~p=0.90$ & 0.458 \\ 
$\hatkmean$ & 0.501 \\ 
$\hat{\mathbf{k}}_p,~p=0.80$ & 0.505 \\ 
$\hat{\mathbf{k}}_p,~p=0.70$ & 0.560 \\ 
$\hatkvar$ & 0.990\\ \hline
\end{tabular}
\label{tab:recap_estimates_metaSWE}
\end{table*}

>>>>

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

@@

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
\usepackage{easy-todo}
\usepackage{graphicx}
\usepackage{booktabs}
% À changer
\newcommand{\yobs}{\mathbf{y}^o}
\DeclareMathOperator*{\argmin}{arg\,min \,}
\DeclareMathOperator*{\argmax}{arg\,max \,}
\newcommand{\Var}{\mathbb{V}\textrm{ar}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Cov}{\textsf{Cov}}
\newcommand{\tra}{\mathrm{tr}}
\newcommand{\kmean}{{\mathbf{k}}_{\mathrm{mean}}}
\newcommand{\hatkmean}{\hat{\mathbf{k}}_{\mathrm{mean}}}
\newcommand{\kvar}{{\mathbf{k}}_{\mathrm{var}}}
\newcommand{\hatkvar}{\hat{\mathbf{k}}_{\mathrm{var}}}
\newcommand{\kmpe}{{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\hatkmpe}{\hat{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\kest}{\hat{\mathbf{k}}}

\newcommand{\checkap}{{\alpha}_p}
\newcommand{\checka}{{\alpha}}
\newcommand{\checkk}{\mathbf{k}}
\newcommand{\checkkp}{{\mathbf{k}}_p}
\newcommand{\Kspace}{\mathbb{K}}
\newcommand{\Uspace}{\mathbb{U}}
\newcommand{\JBH}{J_{\mathrm{BH}}}
\newcommand{\JBHS}{J_{\mathrm{BHswap}}}

\newcommand{\arthur}[1]{{\itshape\color{cyan} ({#1})}}
\newcommand{\elise}[1]{{\itshape\color{red} ({#1})}}
\newcommand{\victor}[1]{{\itshape\color{green} ({#1})}}

@@

Numerical models are widely used to study or forecast natural phenomenon and improve industrial processes. However, by essence models only partially represents reality and sources of uncertainties are ubiquitous (discretisation errors, missing physical processes, poorly known boundary conditions).
%Many physical phenomena are modelled numerically in order to better understand and/or to predict their behaviour. Because of the complexity and or ignorance of the underlying physics, numerical models are imperfect representations of the reality. 
%
%Various sources of uncertainties can corrupt the model quality, such as the discretisation of the domain, or presence of  small scale phenomena that can not be fully represented. 
Moreover, such uncertainties may be of different nature. 
~\cite{walker_defining_2003} propose two to consider two categories of uncertainties. On the one hand aleatoric uncertainties,  coming from the inherent variability of a phenomenon, e.g.\ intrinsic randomness of some environmental variables. On the other hand, epistemic uncertainties coming from a lack of knowledge about the properties and conditions of the phenomena underlying the behaviour of the system under study.
The latter can be accounted for through the introduction of ad-hoc correcting terms in the numerical model, that need to be properly estimated. Thus, reducing the epistemic uncertainty can be done through parameters estimation approaches. 
%when one is having observations of the reality. 
This is usually done using optimal control techniques, leading to an optimisation of a well chosen cost function which is typically built as a comparison with reference observations.
%
An application of such an approach, in the context of ocean circulation modeling, is the estimation of ocean bottom friction parameters in~\cite{das_estimation_1991} and~\cite{boutet_estimation_2015}.

@@

Accounting for the probabilistic nature of  $\mathbf{U}$ leads to study the distribution of the random variable $J(\mathbf{k},\mathbf{U})$, or the distribution of its minimisers ${\mathbf{k}}_{\mathbf{U}}$. The latter is referred as 
the distribution of the conditional minimisers, notion that appeared notably in~\cite{villemonteix_informational_2006} and in~\cite{hennig_entropy_2011} for a global optimisation purpose. Both approaches and related robust estimates are described in Section~\ref{robust formulations}. Section~\ref{sec:relax_constraint} introduces a new class of estimators, by relaxing the constraint of optimality and defining regions of acceptability, similarly as~\cite{buhmann_robust_2013} in discrete combinatorial problems.  The intention is that a robust estimate provides values of the cost function close enough to the attainable minimum for each configuration induced by $\mathbf{u}\in\Uspace$. Illustration of the various discribed methods are given on a numerical exemple in Section~\ref{sec:SWE_application}.

@@

As mentioned before, robustness can be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$. Under this definition, one may design classical robust estimators, either by optimising the moments of the cost function; or based on the distribution of its minimisers.

@@

As the computation of this Pareto front is usually hard and expensive, alternative strategies based on the minimisation of a scalarized version of the vector of objectives are often considered. Some are based on a weighted sum of the objectives, as presented in~\cite{grodzevich_normalization_2006} and in~\cite{marler_weighted_2010}, while some others are based on the minimisation of one of the objectives under constraints on the others, as performed in~\cite{lehman_designing_2004}. Both of these methods are based on an \textit{delicate} choice of weights or of constraints before any computation. This choice relies heavily on a knowledge of the properties of the cost function.

@@

Once again, a set of samples  $\{\mathbf{u}^i\}_{i=1\dots N}$ can be used to compute the set $\{\mathbf{k}_{\mathbf{u}^i}\}_{i=1\dots N}$, from which one can 
approximate  $p_{\mathbf{K}^*}$. The resulting approximation and therefore its mode, is sensitive to the density estimation method. Main methods are KDE  (Kernel Density Estimation)~\cite{elise}, and  EM (Expectation-Maximisation)~\citep{dempster_maximum_1977}.

@@

\begin{figure*}[!ht]
  \centering
  \includegraphics[width=12cm]{Figures/branin_side_moments_noscale.pdf}
  \caption{Left: Contour of the $J_{\mathrm{BH}}$ function, conditional mean and variance. The estimates $\hatkmean$ and $\hatkvar$ are plotted with the dashed line. Right: $J_{\mathrm{BHswap}}$ function. The scale of the bottom plots have not been represented.} %TODO revoir
\label{fig:branin_moments}
\end{figure*}
%\clearpage % Pour lisibilité dans le preview

@@

Illustrations related to $\hatkmpe$ are depicted on Figure~\ref{fig:contours}. As before, the set of $\mathbf{u}^i$ is taken from a regular grid. Top row shows the contour plot of both functions as well as the set of conditional minimisers $\{\mathbf{k}_{\mathbf{u}^i}\}_{1\leq i \leq N}$ in red, as defined in Eq.~\eqref{eq:conditional_minimiser}.
The left, respectively right, column stands for $J_{\mathrm{BH}}$, respectively $J_{\mathrm{BHswap}}$ function. The bottom rows presents three approximations of the density of minimisers: the histogram in grey (bin size selected using Freedman-Diaconis from~\cite{freedman_histogram_1981}), the result of a kernel density estimation (KDE) with Gaussian kernels in red (using Scott's rule from~\cite{scott_optimal_1979} for bandwidth selection), and the estimation by a Gaussian mixture. The latter has been calculated with the EM algorithm. The number of Gaussians has been fixed to 3, a guess based on the general shape of the histogram. Respective estimations of $\hatkmpe$ are also depicted using dashed lines.

@@

\begin{figure*}
  \centering
  \includegraphics[width=12cm]{Figures/branin_side_66.pdf}
  \caption{Top Left:  $J_{\mathrm{BH}}$ function along with conditional minimisers in red. Bottom left: Estimated density using KDE and EM algorithm and conditional mean. The dashed lines indicate the MPE found using those methods. Right: Same quantities with $J_{\mathrm{BHswap}}$. The estimation of the MPE is consistent between the methods for $J_{\mathrm{BHswap}}$, but not for  $J_{\mathrm{BH}}$}
\label{fig:contours}
\end{figure*}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=12cm]{Figures/relaxation_tuto.pdf}
\caption{Principle of the relaxation of the constraint on $J_{\mathrm{BHswap}}$, and illustration of $R_\alpha(\mathbf{k})$}
\label{fig:relax_tuto}
\end{figure*}

%In order to construct this region, we introduce a factor $\alpha \geq 1$, that will be multiplied to $J^*$:
%for a given $\mathbf{u}$, $\mathbf{k}$ is deemed acceptable if $J(\mathbf{k},\mathbf{u}) \leq \alpha J^*( \mathbf{u})$.
Introducing the random nature of $\mathbf{U}$, $\Gamma_\alpha(\mathbf{k})$ is defined as the probability that $\mathbf{k}$ is acceptable given $\alpha$:
\begin{equation}
  \label{eq:def_Gamma}
  \Gamma_\alpha(\mathbf{k}) = \Prob_{\mathbf{U}}\left[\mathbf{U} \in R_\alpha(\mathbf{k})\right] = \Prob_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U}) \leq 
    \alpha J^*(\mathbf{U}) \right]
\end{equation}
In other words, $\Gamma_{\alpha}(\mathbf{k})$ is then the probability that $J(\mathbf{k},\mathbf{U})$ is between 
$J^*(\mathbf{U})$ and 
$\alpha J^*(\mathbf{U})$. Without relaxation, i.e.\ when $\alpha$ is set to $1$, $\Gamma_1$ is non-zero if the set $\{\mathbf{u}\in\Uspace \mid J(\mathbf{k},\mathbf{u}) = J^*(\mathbf{u})\}$ has non-zero measure with respect to $\Prob_{\mathbf{U}}$, that is when the distribution of $\mathbf{K}^*$ presents atoms. This can be linked to the definition of the MPE from Eq.~\eqref{eq:MPE}. In the more general case, the MPE can be seen as the limiting case when $\alpha$ tends to $1$.

@@

Given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$. This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$.

\begin{figure}[!ht]
\centering
\includegraphics[width = 8.3cm]{Figures/illu_alpha_p.pdf}
\caption{Illustration of the influence of different levels $p$ on $\Gamma_{\alpha_p}$ and on $\mathbf{k}_p$ \victor{Pertinence ?}}
\label{fig:illu_alpha_p}
\end{figure}

@@

\begin{table*}[t]
  \centering
\label{tab:RO_recap}
\begin{tabular}{lrr}
  \toprule
  Definition & Related quantities & Interpretation \\ \midrule
   $\argmin_{\mathbf{k}\in\Kspace} \Ex_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kmean$ & Long run performances\\
   $\argmin_{\mathbf{k}\in\Kspace} \Var_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kvar$  & Steady performances\\
   $\argmax_{\mathbf{k}\in\Kspace} p_{\mathbf{K}^*}(\mathbf{k})$ & $\kmpe$ & Most probable minimiser\\
    $\inf\left\{ \alpha \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\}$ & $(p, \checkkp,\checkap)$ & Acceptable values with fixed probability $p$ \\
    $p_{\mathrm{maxratio}} = \argmax p/\checkap$ & $(p_{\mathrm{maxratio}}, \checkk_{\mathrm{maxratio}}, \checka_{\mathrm{maxratio}})$ & Maximal ratio of $p$ and $\checkap$ \\
  \bottomrule
\end{tabular}
  \caption{Robust estimators, based on a cost function $J$}

\end{table*}

@@

% ----------------------------------------------------------------------------
\subsection{Estimation of $\checkkp$ and $\checkap$ for the $J_{\mathrm{BH}}$ and $J_{\mathrm{BHswap}}$ functions}
%Numerical application}
As stated before, we chose to model the uncertainties as a random variable uniformly distributed on $\Uspace$. The bounded nature of $\Uspace$ allow us to consider the estimate introduced before up to a level of confidence $p=1$.
From now on, $\hatkmpe$ is estimated using KDE with Gaussian kernels.
The smallest relaxation $\hat{\checka}_1$ and the corresponding $\hat{\checkk}_1$ has been computed for the $J_{\mathrm{BH}}$ and $J_{\mathrm{BHswap}}$ functions, using a regular grid of $1000 \times 1000$ points on $\Kspace\times\Uspace$. The contour plots of those functions can be seen in the top plots of Figure~\ref{fig:contour_alpha}. The frontier corresponding to the couples of points $(\mathbf{k},\mathbf{u})$ verifying $\{J(\mathbf{k},\mathbf{u}) = \alpha J^*(\mathbf{u})\}$ has been drawn on top of these contour plots, for $\alpha=\hat{\checka}_1$ and an arbitrary $\alpha=1.5<\hat{\checka}_1$, to illustrate the effect of the acceptable region when the relaxation $\alpha$ increases. 
On the bottom plots, the curves $\mathbf{k}\mapsto \hat{\Gamma}_{\alpha}(\mathbf{k})$ for $\alpha=\hat{\checka}_1$ and $\alpha=1.5$ along with the histogram of the minimisers are represented.
One can notice that the relaxation does not present an issue with the accumulation of the minimisers of $\JBH$ at $3.8$, as opposed to the MPE and its dependence on the estimation procedure of the distribution.

@@

\begin{table*}[!ht]
\centering
\caption{Estimation performed for $J_{\mathrm{BH}}$, sorted by value}
\label{tab:recap_estimates_branin}
\begin{tabular}{lr}
\toprule
Estimate & Value \\ \midrule
$\hatkvar$ & 1.371 \\ 
$\kest_p,~p=1$ & 1.557 \\ 
$\hatkmean$ & 1.587 \\ 
  $\hatkmpe$ & 1.628 \\
  $\kest_{\mathrm{maxratio}},~\hat{p}_{\mathrm{maxratio}}=0.654$ & 1.637 \\
$\kest_p,~p=0.95$ & 1.672 \\ \bottomrule
\end{tabular}
\end{table*}

Practically speaking, in order to compare the effective values taken by the objective function given an estimate $\kest$ we are going to consider the functions of the form $\mathbf{u} \mapsto J(\kest,\mathbf{u})$, that we will call ``profile of $\kest$''. Those profiles are well suited for the representation of the cost function for an estimate $\kest$ fixed as the uncertain variable is modelled with a 1D uniform random variable.

@@

\begin{table*}[!h]
\centering
\caption{Estimations performed for $J_{\mathrm{BHswap}}$, sorted by value}
\label{tab:recap_estimates_BHs}
\begin{tabular}{lr} \toprule
Estimate & Value \\ \midrule
$\kest_{\mathrm{MPE}}$ & 0.606 \\ 
$\kest_{\mathrm{maxratio}},~ \hat{p}_{\mathrm{maxratio}}=0.766$ & 1.537 \\ 
$\kest_{\mathrm{mean}}$ & 1.752 \\ 
$\kest_p,~p=0.90$ & 2.112 \\ 
$\kest_p,~p=0.95$ & 2.457 \\ 
$\kest_{\mathrm{var}}$ & 2.638 \\ 
$\kest_1,~p=1$ & 2.798   \\ \bottomrule
\end{tabular}
\end{table*}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=.7\textwidth]{Figures/RobustSWEmis_plots.pdf}
\caption{Procedure of robust calibration for the shallow water problem. Top: contours of $J$, conditional minimisers and $\{(\mathbf{k},\mathbf{u}) | J(\mathbf{k},\mathbf{u}) = \hat{\alpha}_1 J^*(\mathbf{u})\}$. Middle: Conditional moments and histogram and KDE of the conditional minimisers. Bottom: $\hat{\Gamma}_{\alpha_p}$ for different levels $p$}
\label{fig:estimation_swe}
\end{figure}
\begin{table*}[!h]
\centering
\caption{Calibrated values of $\mathbf{k}$ according to different criteria}
\begin{tabular}{lr} \toprule
Estimator & Value \\ \midrule
$\hatkmpe$ & 0.249 \\
$\kest_{\mathrm{global}}$ & 0.290 \\
$\hat{\mathbf{k}}_{p},~p=1$ & 0.423 \\ 
$\hat{\mathbf{k}}_p,~p=0.90$ & 0.458 \\ 
$\hatkmean$ & 0.501 \\ 
$\hat{\mathbf{k}}_p,~p=0.80$ & 0.505 \\ 
$\hat{\mathbf{k}}_p,~p=0.70$ & 0.560 \\ 
$\hatkvar$ & 0.990\\ \bottomrule
\end{tabular}
\label{tab:recap_estimates_metaSWE}
\end{table*}
